---
title: "Working Title: Exploration of Disparities in Healthcare Access in the USA by Race and Location"
author: "Anushka Damidi (damidia@uw.edu), Riley Mintz (rimintz@uw.edu) & Max Stewart (mzs11@uw.edu)"
date: "11/13/2022"
output: html_document
---

```{r setup, include=FALSE}
source("../source/summary_info.R")
source("../source/summary_table.R")
source("../source/chart_1_hospitals.R")
source("../source/chart_2_hospitals_map.R")
source("../source/chart_3_maternal_mortality.R")
library(tidyverse)
library(knitr)
library(ggplot2)
knitr::opts_chunk$set(echo = FALSE)

# dynamic paragraph variables:

```

## Dynamic Paragraph \*pls replace w better title here

In our exploratory analysis, we looked into the information in the three .csv files we've already selected. These statistics are a good start, but we may need to supplement our analysis with other data in order to connect our areas of interest. EXAMPLE: Washington State has `r num_hospitals_in_wa` open hospitals.

## Table: Number of Open Hospitals

```{r fig.align = 'left', echo = FALSE, fig.width=16}
knitr::kable(
  hospital_info,
  format = "simple",
  col.names = c("State","Number Open Hospitals","Owner Unavailable","Owner: Proprietary","Owner: LLC","Owner: Non-Profit","Owner: Federal Govt.","Owner: State Govt.","Owner: District/Authority Govt.","Owner: Local Govt."),
  align = "llllllllll",
  caption = "This table displays the total number of open hospitals in each state, U.S. territory, and D.C., as well as how many of them are owned by each type of owner.",
  )
```

### About this table:

## Plots

```{r fig.align='left', echo=FALSE, fig.width=16}
plot(hospitals_per_state_chart_w_owner)
```

### About this plot

blah blah blah

```{r fig.align='left', echo = FALSE, fig.width=16, fig.height=10}
plot(hospitals_map)
```

### About this plot
We included this chart because our data

```{r echo = FALSE}
plot(maternal_mortality_by_race_chart)
```

### About This Plot



