---
title: "Exploration of Disparities in Healthcare Access in the USA by Race and Location"
author: "Anushka Damidi (damidia@uw.edu), Riley Mintz (rimintz@uw.edu) & Max Stewart (mzs11@uw.edu)"
date: "11/14/2022"
output: html_document
---

```{r setup, include=FALSE}
source("../source/summary_info.R")
source("../source/summary_table.R")
source("../source/chart_1_hospitals.R")
source("../source/chart_2_hospitals_map.R")
source("../source/chart_3_maternal_mortality.R")
library(tidyverse)
library(knitr)
library(ggplot2)
knitr::opts_chunk$set(echo = FALSE)

# dynamic paragraph variables:

```

## Exploratory Analysis Findings
This project will be focusing on the disparities in maternal death and the lack of proper access to healthcare that women from various ethnicities might face. Women of color are marginalized in the healthcare system and this is shown through our sources and the data that was collected around the differing maternal mortality rates based on ethnicity and location. Our sources have data focusing on maternal mortality by race, maternal mortality by location and hospitals by location. Throughout these various data sets, there is a clear distinction between the positive effects of the healthcare system on the different races. For example, there were 16.9 pregnancy related deaths per 100,000 deaths in 2016, 17.3 deaths in 2017 and 17.3 deaths in 2018. However, of those deaths, 41.4 percent were of Non-hispanic Black heritage, and 26.5 percent were of American Indian or Alaskan Native heritage, while only 13.7 percent of those deaths were of White heritage. This shows the drastic difference between maternal mortality rate between the races and how people of certain ethnicities are greatly disadvantaged and are lacking proper care. 

## Table: Number of Open Hospitals

```{r fig.align = 'left', echo = FALSE, fig.width=16}
knitr::kable(
  hospital_info,
  format = "simple",
  col.names = c("State","Number Open Hospitals","Owner Unavailable","Owner: Proprietary","Owner: LLC","Owner: Non-Profit","Owner: Federal Govt.","Owner: State Govt.","Owner: District/Authority Govt.","Owner: Local Govt."),
  align = "llllllllll",
  caption = "This table displays the total number of open hospitals in each state, U.S. territory, and D.C., as well as how many of them are owned by each type of owner.",
  )
```

### About this table
We are interested in seeing how Race and location (within the USA) affect the maternal mortality rate. We will need more data to relate these variables, so for now we are finding patterns in the data we already have. This chart computes the number of hospitals in each US State, territory, and DC, as well as who owns these hospitals. The number of hospitals in each state is relevant because how many hospitals are available in an area is likely to affect the health outcome of the people in that state, such as maternal mortality. We also included who owns the hospitals in each state, because who owns and operates a hospital might affect the quality of care. 


## Plots

```{r fig.align='left', echo=FALSE, fig.width=16}
plot(hospitals_per_state_chart_w_owner)
```

### About this plot
This plot lets viewers see the numbers of open hospitals within each state, as well as who/what owns them. It is a good fit for our data because each state is clearly shown, with its number of hospitals shown on the y-axis and the index explaining the colors that correlate with the owner of the hospital. Using this chart, we can conclude that Texas and California have the highest number of open hospitals in the United States because their bars are much higher than the rest. 

```{r fig.align='left', echo = FALSE, fig.width=16, fig.height=10}
plot(hospitals_map)
```

### About this plot
We included this chart because we wanted a way to easily view the number of open hospitals in each state. A map is the best way to display this information because darker red signifies a higher number of open hospitals and lighter red signifies a lower number of open hospitals. With this chart, we can quickly look and see the highs and lows. For example, Texas has the lowest amount of open hospitals in the country. Note: West Virginia's color coding does not show up on the map regardless of the state name or abbreviation that is written. 

```{r echo = FALSE}
plot(maternal_mortality_by_race_chart)
```

### About This Plot
In this plot, we see how race and ethnicity are related to pregnancy related mortality data. We believe this chart would work well with the data we have wrangled because it shows exactly what is needed to be understood by the viewer. For example, just by glancing at the chart, we can see that non-Hispanic Black people have the highest pregnancy related mortality ratio, and Hispanic people with the lowest. 